### 2015-02-10 Article/Article.tex
09_02_2015

Resource ratio in a given plant interacts with resource ratio of the herbivore, thus constraining persistence of herbivores. It is a bottom-up effect mediated by stoichiometry.
Consider $p$ herbivores feeding on one plant species embedding $k$ resources. According to the Liebig's law of the minimum expressed by the minimum function in equation 1, for each herbivore $i$ the interplay between resource requirement and resource availability will determine which resource is the most limiting for growth \citep{Grover1997}. Depending on their requirements, and on the ratios of resources available in plant biomass, two herbivores feeding on a single plant species may not be limited by the same resource. Thus, requirement differences among herbivores can be interpreted as niche segregation. %Yet, as we shall see, this niche segregation does not allow coexistence when two herbivore species feed on a single plant species.\\ 
For a single herbivore $i$, feeding on the plant, the level of limiting resource $R_u$ (with $u \in [1,k]$) in plant biomass at equilibrium writes (see appendix 3 for details):
\begin{equation}\label{resourceoneplant}
\barre{R}_{uhi}=\barre{V}|_{Ru}\; Q_{vRu} =\frac{m_i}{g_i}Q_{hiRu} 
\end{equation}
Where $\barre{V}|_{Ru}$ is plant biomass at steady state when $R_u$ is limiting. Similarly to cases 1 and 2, this level represents the level of resource available to another herbivore and is therefore a key driver of competitive outcomes. It can be graphically represented by a zero net growth isocline (ZNGI) in the phase space of the $k$ resources in plant biomass. One can also represent how resources are consumed by the herbivore with a consumption vector in this phase space \citep{Tilman1980}. As for the classical resource-ratio theory \citep{Tilman1982}, we will show here that this graphical representation is a valuable tool to address the competitive outcomes between herbivore species. \par
For simplicity, consider two herbivore species ($H_1$ and $H_2$) competing for two resources ($R_1$ and $R_2$) embedded into a single plant species ($V$). According to equation \ref{resourceoneplant}, the slopes of the ZNGIs only depend on herbivore parameters, which are constant. Therefore, the ZNGIs are parallel to the axis on the phase plan $\{R_1,R_2\}$. According to the classic resource-ratio theory \citep{Tilman1982}, a first necessary condition for coexistence at equilibrium is that the ZNGIs of the two herbivores cross one another. A second necessary condition relies on the difference between the slopes of the consumption vectors of the competitors \citep{Tilman1982}.  The consumption vectors graphically illustrate how resource consumption drives the levels of resources from resource supply to equilibrium levels. In the phase plan, resource supply is represented by the supply point $S$, that is, the total amount of resources at equilibrium in absence of consumption:

\begin{equation}
S_{R_1}=Q_{vR_1}(S-a\barre{V})
\end{equation}
\begin{equation}
S_{R_2}=Q_{vR_2}(S-a\barre{V})
\end{equation}
For a given herbivore $i$, the consumption vector $\vec{C_i}$ writes:
\begin{equation}
\vec{C_i}=\left( 
\begin{array}{l}
g_i\barre{V}\barre{H}_i Q_{vR_1}\\
g_i\barre{V}\barre{H}_i Q_{vR_2}
\end{array}\right) 
=g_i\barre{V}\barre{H}_i \left( 
\begin{array}{l}
Q_{vR_1}\\
Q_{vR_2}
\end{array}\right) 
\end{equation}
The vectorÕs slope is determined by the ratios of resources embedded in plant biomass. An important consequence is that all herbivores have the same slope of consumption vector when feeding on a single given plant. This constraint precludes coexistence, and the resource-ratio embedded in plant biomass determines which of the competitor will displace the other (figure \ref{herbifig}). Hence, despite niche segregation on the requirement axe, two herbivores cannot coexist on two resources. This result is in contrast with the classic resource-ratio theory, which stipulates that two consumer may coexist on two resources. The reason is that classically, consumers are assumed to control the slope of their consumption vector \citep{Tilman1982}. This assumption is valid for primary producers, which take up essential nutrients independently in their environment, in the form of dissolved chemical molecules, but not for herbivores, which take up essential nutrients already bounded in plant biomass. Therefore, niche segregation alone cannot promote coexistence. Only herbivore that can match plant stoichiometry the best can persist and outcompete other herbivores.

@@

By looking at boundary ZNGI equation (\ref{boundaryZNGI}), it appears that ZNGI slopes depends only on herbivore constant parameters. Therefore, it is parallel to the axis. Moreover, two herbivores having similar quotas or different feeding strategies would have different ZNGI. Thus, if boundary ZNGI of two herbivores cross each other, coexistence might be possible. \par

Next, competition outcome will depend on the relative orientation of the consumption vector of herbivores \citep{Tilman1980}. However, vector slope calculation is uneasy in the case of packaged resources because vector slope depends on plant supply points. Hence, for a given herbivore, vectors do not have a constant slope valid for all supply conditions. This is why instead of consumption vectors, the model allows the calculation of boundary vectors that determine an area in the phase plan where each herbivore species could coexist with an other one. It is possible to determine a boundary relation between plant supply points ($S_1$ and $S_2$) that allows each herbivore to consume both resources when one resource is limiting, and leads to $\barre{V}_1$ and $\barre{V}_2$ at steady state. These boundary supply points are $S_{1hi}$ and $S_{2hi}$ for plant 1 and 2 respectively (see appendix 4). Hence, boundary vector slopes are:
\begin{equation}
\vec{C}_{h1}= 
\begin{pmatrix}
\barre{S}_{1h1} \;Q_{v1R1}+\barre{S}_{2h1}\; Q_{v2R1} \\
\barre{S}_{1h1}\; Q_{v1R2}+\barre{S}_{2h1}\; Q_{v2R2} \\
\end{pmatrix} 
\end{equation}
\begin{equation}
\vec{C}_{h2}= 
\begin{pmatrix}
\barre{S}_{1h2} \;Q_{v1R1}+\barre{S}_{2h2}\; Q_{v2R1} \\
\barre{S}_{1h2}\; Q_{v1R2}+\barre{S}_{2h2}\; Q_{v2R2} \\
\end{pmatrix} 
\end{equation}
These two boundary vectors play a similar role as consumption vectors in Tilman's model. However, as resources are packaged into plants, vector slopes depend on the way herbivore species feed on plants containing resources.  
Hence, assuming plant 1 is richer in resource 1 ($Q_{v1R1}>Q_{v2R1}$), and plant 2 is richer in resource 2 ($Q_{v1R2}<Q_{v2R2}$), and assuming herbivore 1 is more limited by resource 1 ($Q_{h1R1}>Q_{h1R2}$), and herbivore 2 is more limited by resource 2 ($Q_{h2R1}<Q_{h2R2}$), three different strategies can be considered. The first one occurs when each herbivore species consumes preferentially the plant species which gives the greater quantity of the most limiting resource for this herbivore species, and which is less limiting for its competitor (i.e., $g_{11}>g_{12}$ and $g_{21}<g_{22}$). In that case, if we consider the boundary vectors for herbivore 1 and for herbivore 2, we can define several zones on the phase plan (see figure \ref{Coexistence}). The zone between the two boundary vectors represents the supply conditions allowing stable coexistence of the two herbivore species. Thus, each herbivore consumes the most profitable plant (i.e., quotas and consumption functions follow a similar trend), but each herbivore is specialized (more or less) on a different plant than its competitor, which promotes coexistence. \par
The second case represents the reverse situation, where each herbivore species consumes preferentially the plant species which provides greater quantity of the most limiting resource for its competitor (i.e., $g_{11}<g_{12}$ and $g_{21}>g_{22}$). Hence, the zone between the boundary vectors does not allow coexistence (see figure \ref{Exclusion}), and in many cases equation \ref{Systeme} does not have any realistic solution (with both herbivore species persisting). Thus, competitive exclusion is the general outcome. \par 
The third case occurs when herbivore species are not selective at all: they consume plant species according to their respective biomass. In that case, both herbivores consume resources in the same way (i.e., $g_{11}=g_{21}$ and $g_{12}=g_{22}$). Hence, there is no segregation in herbivore consumption: all herbivore species have the same consumption vector, which is similar to case 3, where only one plant species is present (see figure \ref{herbifig}). Therefore, the trajectory will either cross herbivore 1 ZNGI first, and this herbivore will be excluded, or the trajectory will cross herbivore 2 ZNGI first, and this herbivore will be excluded. Only a particular case, where the trajectory leads exactly on the crossing point, allows coexistence. \par
If we release the previous assumptions on herbivore and plant quotas (i.e., inequality for herbivore requirements, and inequality for plant profitability), two other situations can potentially occur. First, both herbivores can have the same quotas (i.e., $Q_{h1R1}=Q_{h2R1}$ and $Q_{h1R2}=Q_{h2R2}$) but different consumption functions. In that case, one plant will be more interesting for both herbivores (i.e., the plant that is richer for the most limiting resource). Hence, the herbivore that can consume this plant the most efficiently will exclude its competitor. Second, both plants can have the same quotas (i.e., $Q_{v1R1}=Q_{v2R1}$ and $Q_{v1R2}=Q_{v2R2}$). In that case, the system is similar to the case where only one plant is present (case 3), which leads to competitive exclusion. \par
Last, due to resource packaging into plants, the supply point is constrained. Resource ratio within each plant will define a slope:
\begin{equation}
\alpha _{vj} = \frac{Q_{vjR2}}{Q_{vjR1}}
\end{equation}
In case of two plants, supply point will lie between the two extreme slopes ($\alpha _{v1}$ and $\alpha _{v2}$) defining a feasibility cone. Therefore, the whole phase plan is not available. Resource supply occurs only within this feasibility cone, which adds another constraint on herbivore persistence and competition.

>>>>

Resource ratio in a given plant interacts with resource ratio of the herbivore, thus constraining persistence of herbivores. It is a bottom-up effect mediated by stoichiometry.
Consider $p$ herbivores feeding on one plant species embedding $k$ resources. According to the Liebig's law of the minimum expressed by the minimum function in equation 1, for each herbivore $i$ the interplay between resource requirement and resource availability will determine which resource is the most limiting for growth \citep{Grover1997}. Depending on their requirements, and on the ratios of resources available in plant biomass, two herbivores feeding on a single plant species may not be limited by the same resource. Thus, requirement differences among herbivores can be interpreted as niche segregation. %Yet, as we shall see, this niche segregation does not allow coexistence when two herbivore species feed on a single plant species.\\ 
For a single herbivore $i$, feeding on the plant, the level of limiting resource $R_u$ (with $u \in [1,k]$) in plant biomass at equilibrium writes (see appendix 3 for details):
\begin{equation}\label{resourceoneplant}
\barre{R}_{uhi}=\barre{V}|_{Ru}\; Q_{vRu} =\frac{m_i}{g_i}Q_{hiRu} 
\end{equation}
Where $\barre{V}|_{Ru}$ is plant biomass at steady state when $R_u$ is limiting. Similarly to former cases, this level represents the level of resource available to another herbivore and is therefore a key driver of competitive outcomes. It can be graphically represented by a zero net growth isocline (ZNGI) in the phase space of the $k$ resources in plant biomass. One can also represent how resources are consumed by the herbivore with a consumption vector in this phase space \citep{Tilman1980}. As for the classical resource-ratio theory \citep{Tilman1982}, we will show here that this graphical representation is a valuable tool to address the competitive outcomes between herbivore species. \par
For simplicity, consider two herbivore species ($H_1$ and $H_2$) competing for two resources ($R_1$ and $R_2$) embedded into a single plant species ($V$). According to equation \ref{resourceoneplant}, the slopes of the ZNGIs only depend on herbivore parameters, which are constant. Therefore, the ZNGIs are parallel to the axis on the phase plan $\{R_1,R_2\}$. According to the classic resource-ratio theory \citep{Tilman1982}, a first necessary condition for coexistence at equilibrium is that the ZNGIs of the two herbivores cross one another. A second necessary condition relies on the difference between the slopes of the consumption vectors of the competitors \citep{Tilman1982}.  The consumption vectors graphically illustrate how resource consumption drives the levels of resources from resource supply to equilibrium levels. In the phase plan, resource supply is represented by the supply point $S$, that is, the total amount of resources at equilibrium in absence of consumption:

\begin{equation}
S_{R_1}=Q_{vR_1}(S-a\barre{V})
\end{equation}
\begin{equation}
S_{R_2}=Q_{vR_2}(S-a\barre{V})
\end{equation}
For a given herbivore $i$, the consumption vector $\vec{C_i}$ writes:
\begin{equation}
\vec{C_i}=\left( 
\begin{array}{l}
g_i\barre{V}\barre{H}_i Q_{vR_1}\\
g_i\barre{V}\barre{H}_i Q_{vR_2}
\end{array}\right) 
=g_i\barre{V}\barre{H}_i \left( 
\begin{array}{l}
Q_{vR_1}\\
Q_{vR_2}
\end{array}\right) 
\end{equation}
The vectorÕs slope is determined by the ratios of resources embedded in plant biomass. An important consequence is that all herbivores have the same slope of consumption vector when feeding on a single given plant. This constraint precludes coexistence, and the resource-ratio embedded in plant biomass determines which of the competitor will displace the other (figure \ref{herbifig}). Hence, despite niche segregation on the requirement axe, two herbivores cannot coexist on two resources. This result is in contrast with the classic resource-ratio theory, which stipulates that two consumer may coexist on two resources. The reason is that classically, consumers are assumed to control the slope of their consumption vector \citep{Tilman1982}. This assumption is valid for primary producers, which take up essential nutrients independently in their environment, in the form of dissolved chemical molecules, but not for herbivores, which take up essential nutrients already bounded in plant biomass. Therefore, niche segregation alone cannot promote coexistence. Only herbivore that can match plant stoichiometry the best can persist and outcompete other herbivores. This leads to a strong bottom-up stoichiometric constraint.

@@

By looking at boundary ZNGI equation (\ref{boundaryZNGI}), it appears that ZNGI slopes depends only on herbivore constant parameters. Therefore, it is parallel to the axis. Moreover, two herbivores having similar quotas but different feeding strategies would have different ZNGI. Thus, if boundary ZNGI of two herbivores cross each other, coexistence might be possible. \par

Next, competition outcome will depend on the relative orientation of the consumption vector of herbivores \citep{Tilman1980}. However, vector slope calculation is uneasy in the case of packaged resources because vector slope depends on plant supply points. Hence, for a given herbivore, vectors do not have a constant slope valid for all supply conditions. This is why instead of consumption vectors, the model allows the calculation of boundary vectors that determine an area in the phase plan where each herbivore species could coexist with an other one. It is possible to determine a boundary relation between plant supply points ($S_1$ and $S_2$) that allows each herbivore to consume both resources when one resource is limiting, and leads to $\barre{V}_1$ and $\barre{V}_2$ at steady state. These boundary supply points are $S_{1hi}$ and $S_{2hi}$ for plant 1 and 2 respectively (see appendix 4). Hence, boundary vector slopes are:
\begin{equation}
\vec{C}_{h1}= 
\begin{pmatrix}
\barre{S}_{1h1} \;Q_{v1R1}+\barre{S}_{2h1}\; Q_{v2R1} \\
\barre{S}_{1h1}\; Q_{v1R2}+\barre{S}_{2h1}\; Q_{v2R2} \\
\end{pmatrix} 
\end{equation}
\begin{equation}
\vec{C}_{h2}= 
\begin{pmatrix}
\barre{S}_{1h2} \;Q_{v1R1}+\barre{S}_{2h2}\; Q_{v2R1} \\
\barre{S}_{1h2}\; Q_{v1R2}+\barre{S}_{2h2}\; Q_{v2R2} \\
\end{pmatrix} 
\end{equation}
These two boundary vectors play a similar role as consumption vectors in Tilman's model. However, as resources are packaged into plants, vector slopes depend on the way herbivore species feed on plants containing resources.  
Hence, assuming plant 1 is richer in resource 1 ($Q_{v1R1}>Q_{v2R1}$), and plant 2 is richer in resource 2 ($Q_{v1R2}<Q_{v2R2}$), and assuming herbivore 1 is more limited by resource 1 ($Q_{h1R1}>Q_{h1R2}$), and herbivore 2 is more limited by resource 2 ($Q_{h2R1}<Q_{h2R2}$), three different strategies can be considered. The first one occurs when each herbivore species consumes preferentially the plant species which gives the greater quantity of the most limiting resource for this herbivore species, and which is less limiting for its competitor (i.e., $g_{11}>g_{12}$ and $g_{21}<g_{22}$). In that case, if we consider the boundary vectors for herbivore 1 and for herbivore 2, we can define several zones on the phase plan (see figure \ref{Coexistence}). The zone between the two boundary vectors represents the supply conditions allowing stable coexistence of the two herbivore species. Thus, each herbivore consumes the most profitable plant (i.e., quotas and consumption functions follow a similar trend), but each herbivore is specialized (more or less) on a different plant than its competitor, which promotes coexistence. \par
The second case represents the reverse situation, where each herbivore species consumes preferentially the plant species which provides greater quantity of the most limiting resource for its competitor (i.e., $g_{11}<g_{12}$ and $g_{21}>g_{22}$). Hence, the zone between the boundary vectors does not allow coexistence (see figure \ref{Exclusion}), and in many cases equation \ref{Systeme} does not have any realistic solution (with both herbivore species persisting). Thus, competitive exclusion is the general outcome. \par 
One can notice that plant stoichiometry constrains the supply point because of resource packaging into plants. 
%Last, due to resource packaging into plants, the supply point is constrained. 
Resource ratio within each plant will define a slope:
\begin{equation}
\alpha _{vj} = \frac{Q_{vjR2}}{Q_{vjR1}}
\end{equation}
In case of two plants, supply point will lie between the two extreme slopes ($\alpha _{v1}$ and $\alpha _{v2}$) defining a feasibility cone. Therefore, the whole phase plan is not available. Resource supply occurs only within this feasibility cone, which adds another constraint on herbivore persistence and competition.

Spatial segregation of resources embedded in plants and bottom-up stoichiometry are the two key points. If resources are not spatially decoupled or if herbivores do not have a feeding strategy that can compensate for the bottom-up stoichiometry imbalance, coexistence is impossible between herbivore species. The following cases illustrate this argument. 
A first case occurs when herbivore species are not selective at all: they consume plant species according to their respective biomass. In that case, both herbivores consume resources in the same way (i.e., $g_{11}=g_{21}$ and $g_{12}=g_{22}$). Hence, there is no segregation in herbivore consumption: all herbivore species have the same consumption vector, which is similar to the case where only one plant species is present (see figure \ref{herbifig}). Therefore, the trajectory will either cross herbivore 1 ZNGI first, and this herbivore will be excluded, or the trajectory will cross herbivore 2 ZNGI first, and this herbivore will be excluded. %Only a particular case, where the trajectory leads exactly on the crossing point, allows coexistence. \par
Coexistence will be unlikely because even if resources are spatially segregated, herbivores are constrained by plant stoichiometry. \par
%If we release the previous assumptions on herbivore and plant quotas (i.e., inequality for herbivore requirements, and inequality for plant profitability), two other situations can potentially occur.
A second case occurs when the previous assumptions on herbivore and plant quotas (i.e., inequality for herbivore requirements, and inequality for plant profitability) are released. %, which leads to two other situations.% can potentially occur.
Therefore, both herbivores can have the same quotas (i.e., $Q_{h1R1}=Q_{h2R1}$ and $Q_{h1R2}=Q_{h2R2}$) but different consumption functions. In that case, one plant will be more interesting for both herbivores (i.e., the plant that is richer for the most limiting resource). Hence, the herbivore that can consume this plant the most efficiently will exclude its competitor. Again, bottom-up stoichiometry constrained herbivore coexistence. 
Another case can occur when both plants have the same quotas (i.e., $Q_{v1R1}=Q_{v2R1}$ and $Q_{v1R2}=Q_{v2R2}$). In that case, the system is similar to the case where only one plant is present, which leads to competitive exclusion. Here, spatial segregation does not occur. \par
